---
title: "Fight for Human Dignity"
author: "Risa Nabari"
date: '2022-11-19'
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/a4-helpers.R")
source("../source/analysis.R")
```

## Abstract 
This report uses the incarceration data set, and the key goal of this data set is to help understand the inequalities associated with people who have been incarcerated and who are in jail, specifically in the US society. We will be analyzing the number of jailed population of people with different colors as well as the growth of the U.S. prison population and state. 

# Test (TODO: Remove before submission!!)
```{r echo=FALSE}
q1_string <- test_query1()
num <- 5
```
As you know, this is a markdown report. The [R Markdown Cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf) is **useful!**

* Here's a test of a dynamic variable. Say **"`r q1_string`!!!** 
* Here's another test. A vector
of numbers: `r test_query2(num)`.

Recall that _{r echo=TRUE}_ can be used to display the code, 
which is very useful when developing and debugging a report. Here's an example: 

```{r echo=TRUE}
num <- 5
sum <- sum(test_query2(num))
print(sum)
```

## Introduction 

###### Understanding the inequalities associated with people who have been incarcerated and who are in jail, specifically in the US society is very crucial in order to eventually fight for human dignity, advocate for immigrants, invest in the communities, and to end mass incarceration. Mass incarceration robs people of dignity and tears at the fabric of communities. The biggest direct stakeholder includes those who are directly affected by this (mostly the people of color sent to jail and their family). The indirect stakeholders includes the researchers, organizers, and the government (leader) who analyzes the data and to create evidence-based solutions that restore communities in order to end criminalization of people of color, immigrants, and people experiencing poverty. If we are able to analyze the data and draw important conclusions and if we are able to work out the solutions, we could create a new paradigm for public safety and racial justice. Thus in this report, we will be analyzing the number of jailed population of people with different colors as well as the growth of the U.S. prison population and state. 

********“Black men comprise about 13% of the population, but 35% of the incarcerated”?**************************************


## Data Summary 
In order to look at the inequalities associated with people who have been incarcerated and who are in jail, we will be looking at the jailed population by race and compare it with the population of each race to find the ratio. This will allow us to compare between each race and discuss if we are able to see a evidence of inequality. I have used the data from 2018 as having the most recent year would be the most relevant. 
The answers I will be answers is :
What is the average value of the jailed population of Asian Americanage 15 to 64?
What is the average value of the jailed population of Black from age 15 to 64?
What is the average value of the jailed population of African American from age 15 to 64?
What are the average value of the jailed population of White from age 15 to 64?
What are the average value of the jailed population of Latinx from age 15 to 64?

First, I have extracted the population by race who are jailed and the total mean population.

Here are the result:

American Asian Pacific Islander : `r overall$ratio_aapi_2018`
Black : `overall$ratio_black_2018`
Latinx :`overall$ratio_latinx_2018`
Native Americans : `overall$ratio_native_2018`
White : `overall$ratio_white_2018`

(To get this result, I have selected the relevant columns, then found the mean jail population by filterting 
in year 2018. Then, found the ratio by dividing mean jail population of the certain race to the mean population from 15 to 64) 

##Section 3 : Growth of the U.S. Prison Population
``` {r Graph 1, echo = FLASE, fig.cap = "caption"} 
plot_jail_pop_for_us()
```

##Section 4 : Jail Population by States
``` {r Graph 2, echo = FLASE, fig.cap = "caption"} 
plot_jail_pop_by_states()
```
##Section 5: Variable Comparison
``` {r Graph e, echo = FLASE, fig.cap = "caption"} 
section_five_graph()
```

Here, we will be having a variable comparison that reveals potential patterns of inequality. 
In section 2, we have revealed the ratio of the mean jailed population to the population of 
various race specific to 2018. We have learned that...

Now, is there a difference between women and men? Which gender may be potentially 
more vulnerable to getting incarceranated in certain states?



In section 5, we will be looking at the trend of the ration of jailed 
population over time. We will be answering: Are the jailed population of the races increaseing or decreasing?
Does help to support the claim of inequality, or does it not? 
--
##Section 6: 
``` {r Graph 4, echo = FLASE, fig.cap = "caption"} 
section_five_graph ()
```

Section 5: <variable comparison that reveals potential patterns of inequality>
Section 3 and section 4 show prison population trends by time. But, notice that these charts do not show patterns of inequality. They do not, for example, answer these kinds of questions:

Are some races (or genders) disproportionally represented in jail in some areas of the country. Note the importance of state population(s).
Do some kinds of counties seem to imprison more people than other kinds of counties? (Note the importance of county population and perhaps other features of a county.)
In this section, your goal is to produce a chart that reveals a potential inequality. This chart should show how two different continuous variables are related to one another. Think carefully about what such a comparison means, and what you want to communicate to your user. Your first step should be to find potential trends in the dataset.

For the report, include the following in this section: 

Chart. Present the chart.
Chart caption. Include a caption that names and briefly describes the chart.
Summary paragraph. Include the following: 
A specific question(s).
A succinct answer to the question, referring to evidence in the chart.
A brief paragraph (50 words or more) that summarizes the key patterns that appear to be revealed in the chart.
Structuring your code. As in the previous sections, you should write two functions (or more, if useful), one for data wrangling and one for plotting.


_All the very best with your work!_

