---
title: "Revealing the Inequality (using incarceration data)"
author: "Risa Nabari"
date: '2022-11-19'
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/analysis.R")
```

## Abstract 
This report uses the incarceration data set, and the key goal of this data set is to help understand the inequalities associated with people who have been incarcerated and who are in jail, specifically in the US society. We will be analyzing the number of jailed population of people with different colors as well as the growth of the U.S. prison population and state. 


## Introduction 

Understanding the inequalities associated with people who have been incarcerated and who are in jail, specifically in the US society is very crucial in order to eventually fight for human dignity, advocate for immigrants, invest in the communities, and to end mass incarceration. Mass incarceration robs people of dignity and tears at the fabric of communities. The biggest direct stakeholder includes those who are directly affected by this (mostly the people of color sent to jail and their family). The indirect stakeholders include the researchers, organizers, and the government (leader) who analyzes the data and to create evidence-based solutions that restore communities in order to end criminalization of people of color, immigrants, and people experiencing poverty. If we are able to analyze the data and draw important conclusions and if we are able to work out the solutions, we could create a new paradigm for public safety and racial justice. Thus in this report, we will be analyzing the number of jailed people with different colors, gender, as well as the growth of the U.S. prison population and state. 

## Data Summary 
In order to look at the inequalities associated with people who have been incarcerated and who are in jail, we will be looking at the jailed population by race and compare it with the population of each race to find the ratio. This will allow us to compare between each race and discuss if we are able to see a evidence of inequality. I have used the data from 2018 as having the most recent year would be the most relevant. 
The answers I will be answers is :  What is the average value of the Asian American jailed population by the mean population of Asian American from age 15 to 64 ? What is the average value of the Black jailed population by the mean population of Black from age 15 to 64?
What is the average value of the African American jailed population by the mean population of African American from age 15 to 64? What are the average value of the White jailed population by the mean population of White from age 15 to 64? What are the average value of the Latinx jailed population by the mean population of Latinx from age 15 to 64?

Here are the result:

#### **American Asian Pacific Islander : `r overall$ratio_aapi_2018`**

#### **Black : `r overall$ratio_black_2018`**

#### **Latinx :`r overall$ratio_latinx_2018`**

#### **Native Americans : `r overall$ratio_native_2018`**

#### **White : `r overall$ratio_white_2018`**

_To get this result, I have selected the relevant columns, then found the mean jail population by filtering in year 2018. Then, found the ratio by dividing mean jail population of the certain race to the mean population from 15 to 64_

We can see that the ratio of black population (and race with color) is considerably higher than the white population. 

## Growth of the U.S. Prison Population
``` {r Graph 1, echo = FALSE, fig.cap = "This graph shows the growth of U.S. prison population from 1970-2018"} 
plot_jail_pop_for_us()
```
Summary paragraph. A brief paragraph (50 words or more) that summarizes the key patterns that appear to be revealed in the chart.

**Summary**: What pattern can we see in the growth of jailed population in U.S. over time?
Overall, we can say that the jail population in the U.S. has increased drastically. From 1970 to 1980, it was pretty consistent where the value was around 180,000. However from 1980 to 2010, it has increased steadily with a steep slope reaching near to around 800,000 in 2008. From there, it has decreased slightly but compared to back in 1970, it has still been high. 

## Jail Population by States
``` {r Graph 2, echo = FALSE, fig.cap = "Growth of prison population in Washington, California, and Oregon over time (from 1970 to 2018)."} 
plot1
```
**Summary**: In all three states, we can see that the prison population has increased over time. Oregon has the least number of prison population and has a slow steady slope. Washington also has a slow steady slope by slightly higher than the prison population of Oregon. On the other hand, California has a high jail population as seen in the graph. From 1970 to 1980, it was steady (although higher than the other two states) but from 1980 to 1990, it had a very steep slope which increased from around 28000 to 70000. Then, the numbers increased and decreased until 2020. But overall, the total jail population has increased. This could be because California is densely populated which can cause disparity between the community. I have chosen these states as I am currently living in Washington so I was curious on what the pattern was like. In addition, I wanted to compare the difference in the states located on the west coast to see if it would give a similar result. 

## Variable Comparison

Here, we will be having a variable comparison that reveals potential patterns of inequality. In data summary, we have revealed the ratio of the mean jailed population to the whole population of different races. Now, is there a difference between women and men? Which gender may be potentially 
more vulnerable to getting incarcerated? 
``` {r Graph e, echo = FALSE, fig.cap = "Variable comparison of jailed population by gender (men and women) in 2018 by states"} 
section_five_graph()
```

**Summary**: Shown in the graph above, we can see that men has a higher number of jailed population compared to women. We can see this easily by the plots shown. For example, the value on the right corner has around 2200 jailed women and the mail population has around 15000, which is a big difference. This tells that overall, more men is in jail and this could potentially relate to inequality, where men are over 8 times more likely than women to be incarcerated in prison. Throughout this report, we have known that people with color and who are men has a higher population to be incarcerated in prison. 

## Potential patterns of inequality that vary geographically
``` {r Graph 4, echo = FALSE, fig.cap = "Presents the jailed black population in the United States (2018) by states. The more darker the color is, the more jailed black population in that state."} 
map_one
```

Section 6: <a map shows potential patterns of inequality that vary geographically>
  In this section, your goal is to produce a map that reveals a potential inequality.
Specifically, the map should show show show how a variable is distributed geographically. 
Again, think carefully about how a "geographic comparison" (e.g., counties in a state, counties in division, or counties in across regions) might reveal an inequality. 
Your first step should be to find potential trends in the dataset.  Recommendation: See reading on maps and (1)  Use a map based coordinate system to set the aspect ratio of your map; 
and (2) Use a minimalist theme for the map (see reading). 

Structuring your code. As in the previous sections, you should write two functions (or more, if useful), one for data wrangling and one for plotting.



--
Section 5: <variable comparison that reveals potential patterns of inequality>
Section 3 and section 4 show prison population trends by time. But, notice that these charts do not show patterns of inequality. They do not, for example, answer these kinds of questions:

Are some races (or genders) disproportionally represented in jail in some areas of the country. Note the importance of state population(s).
Do some kinds of counties seem to imprison more people than other kinds of counties? (Note the importance of county population and perhaps other features of a county.)
In this section, your goal is to produce a chart that reveals a potential inequality. This chart should show how two different continuous variables are related to one another. Think carefully about what such a comparison means, and what you want to communicate to your user. Your first step should be to find potential trends in the dataset.

For the report, include the following in this section: 

Chart. Present the chart.
Chart caption. Include a caption that names and briefly describes the chart.
Summary paragraph. Include the following: 
A specific question(s).
A succinct answer to the question, referring to evidence in the chart.
A brief paragraph (50 words or more) that summarizes the key patterns that appear to be revealed in the chart.
Structuring your code. As in the previous sections, you should write two functions (or more, if useful), one for data wrangling and one for plotting.