---
title: "Fight for Human Dignity"
author: "Risa Nabari"
date: '2022-11-19'
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/a4-helpers.R")
source("../source/analysis.R")
```

## Abstract 
This report uses the incarceration data set, and the key goal of this data set is to help understand the inequalities associated with people who have been incarcerated and who are in jail, specifically in the US society. We will be analyzing the number of jailed population of people with different colors as well as the growth of the U.S. prison population and state. 

# Test (TODO: Remove before submission!!)
```{r echo=FALSE}
q1_string <- test_query1()
num <- 5
```
As you know, this is a markdown report. The [R Markdown Cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf) is **useful!**

* Here's a test of a dynamic variable. Say **"`r q1_string`!!!** 
* Here's another test. A vector
of numbers: `r test_query2(num)`.

Recall that _{r echo=TRUE}_ can be used to display the code, 
which is very useful when developing and debugging a report. Here's an example: 

```{r echo=TRUE}
num <- 5
sum <- sum(test_query2(num))
print(sum)
```

## Introduction 

###### Understanding the inequalities associated with people who have been incarcerated and who are in jail, specifically in the US society is very crucial in order to eventually fight for human dignity, advocate for immigrants, invest in the communities, and to end mass incarceration. Mass incarceration robs people of dignity and tears at the fabric of communities. The biggest direct stakeholder includes those who are directly affected by this (mostly the people of color sent to jail and their family). The indirect stakeholders includes the researchers, organizers, and the government (leader) who analyzes the data and to create evidence-based solutions that restore communities in order to end criminalization of people of color, immigrants, and people experiencing poverty. If we are able to analyze the data and draw important conclusions and if we are able to work out the solutions, we could create a new paradigm for public safety and racial justice. Thus in this report, we will be analyzing the number of jailed population of people with different colors as well as the growth of the U.S. prison population and state. 

********“Black men comprise about 13% of the population, but 35% of the incarcerated”?**************************************


## Data Summary 
In order to look at the inequalities associated with people who have been incarcerated and who are in jail, we will be looking at the jailed population by race and compare it with the population of each race to find the ratio. This will allow us to compare between each race and discuss if we are able to see a evidence of inequality. 

First, I have extracted the population by race who are jailed and the total mean population.

Here are the result:
  
  
What is the average value of the jailed population of black or African American from age 15 to 64?
What are the average value of the jailed population of white from age 15 to 64?
What are the average value of the jailed population of latino from age 15 to 64?

``` {r Graph 1, echo = FLASE, fig.cap = "caption"} 
plot_jail_pop_for_us()
```


``` {r Graph 2, echo = FLASE, fig.cap = "caption"} 
plot_jail_pop_by_states()
```

``` {r Graph 3, echo = FLASE, fig.cap = "caption"} 
plot_jail_pop_by_states()
```
[TODO]

[TODO ... other sections ... See Canvas for details]

_All the very best with your work!_

